---
subtitle: "Data Visualisation with R"
pagetitle: "Visualising multivariate data with ggplot2"
author: "Emi Tanaka"
date: "28th November 2022"
host: "Australasian Applied Statistics Conference 2022"
institute: "Monash University"
strip-comments: true
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: true
    theme: assets/monash.scss
    show-slide-number: all
    controls: true
    width: 1280
    height: 720
    incremental: true
    auto-stretch: false
    css: [assets/custom.css, assets/pacman.css, assets/tachyons-addon.css]
    include-after: |
      <link rel="stylesheet" href="assets/syntax-highlight.css">
      <link rel="stylesheet" href="assets/fontawesome-free-6.1.1-web/css/all.min.css">
      <script src="assets/fontawesome-free-6.1.1-web/js/all.min.js"></script>
callout-icon: false
execute:
  echo: true
  cache: true
---


## {#title-slide background-image="images/bg5.jpeg"}

```{r}
#| label: setup
#| include: false
library(tidyverse)
source("setup.R")
```

```{r}
#| child: title-slide.qmd
```



## Examining two or more variables

```{r}
#| echo: false
#| output: asis
fns <- dir(here::here("slides/images/catalogue/")) %>% 
  str_subset("(^corr|^other)")
cat(paste0("<img class='catalogue' src='images/catalogue/", fns, "'>", collapse = ""))
```

## Illustrative data 💎 diamonds

* Let's have a look at the `diamonds` data which contains information about different attributes of diamonds

```{r}
data(diamonds, package = "ggplot2")
diamonds
```

## A scatterplot with `geom_point()`

```{r geom-point}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_point()
```

* Scatter plot of price vs carat of diamonds
* Each point corresponds to a diamaond

::: notes

* A scatterplot is useful in visualising two numerical variables 
* Each point corresponds to an observational unit 
* If you have other variables, you may like to map this to other aesthetic values, such as `shape` or `color`.
* When there is a lot of observations, it may be useful to use `alpha` to control the transparency of the points. 

:::


## A hexagonal 2D heatmap with `geom_hex()`

```{r geom-hex}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_hex()
```

* A hexagon shows the count of observations within the region.

::: notes

* A hex plot is useful if there is a lot of _overplotting_ in your scatterplot.
* **_Overplotting_** in data visualisation means that the data or labels overlap in the data visualisation making it harder to distinguish individual data points. * Overplotting typically occurs when you have a large dataset.
* A hex plot overcomes the issue by showing the number of observations within a hexagonal area.
* The count is shown by the `fill` in the resulting data visualisation.

:::


## A rectangular 2D heatmap with `geom_bin_2d()`

```{r geom-bin2d}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_bin_2d()
```


::: notes

* `geom_bin_2d()` is the same as `geom_hex()` but the shape is a rectangle instead of a hexagon.

:::


## Count of overlapping points with `geom_count()`

```{r geom-count}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_count()
```

::: notes

* `geom_count()` counts the number of points in the same location
* The count is then mapped to the aesthetic `size`. 
:::


## Contour of a 2D density with `geom_density_2d()`

```{r geom-density-2d}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_density_2d()
```

::: notes

* If you have want to visualise the joint density of two variables then `geom_density_2d()` is useful. 
* The axis shows the density of the corresponding variable.
* The 2D density is estimated using `MASS::kde2d()`. 

:::


## Contour of a 2D density with `geom_density_2d_filled()`

```{r geom-density-2d-filled}
ggplot(data = diamonds,
       mapping = aes(x = carat, y = price)) +
  geom_density_2d_filled()
```

::: notes

* `geom_density_2d()` draws the contour line 
* `geom_density_2d_filled()` draws the contour bands

:::


## Illustrative data ⛰️ volcano 

* The data contains topographic information on Auckland's Maunga Whau Volcano

```{r}
volcanod <- data.frame(row = as.vector(row(volcano)),
                       col = as.vector(col(volcano)),
                       value = as.vector(volcano))

volcanod
```

::: notes

* The original data is a matrix and we must wrangle in the form ready for `ggplot` first.

:::


## A 2D contour plot with `geom_contour()`

```{r geom-contour}
ggplot(data = volcanod, 
       mapping = aes(x = col, y = row, z = value)) + 
  geom_contour()
```



## A 2D contour plot with `geom_contour_filled()`

```{r geom-contour-filled}
ggplot(data = volcanod, 
       mapping = aes(x = col, y = row, z = value)) + 
  geom_contour_filled()
```




## Tile plots with `geom_tile()`

```{r geom-tile}
ggplot(volcanod, aes(col, row, fill = value)) +
 geom_tile()
```


## Raster plots with `geom_raster()`

```{r geom-raster}
ggplot(volcanod, aes(col, row, fill = value)) +
 geom_raster()
```

* A high performance special case of `geom_tile()` for when the tiles are the same size


## Rectangular plots with `geom_rect()`

```{r geom-rect}
ggplot(volcanod, aes(xmin = col - 0.5, xmax = col + 0.5,
                     ymin = row - 0.5, ymax = row + 0.5, fill = value)) +
 geom_rect()
```

* A reparameterised version of `geom_tile()`


## Illustrative data 📈 economics


```{r}
data(economics, package = "ggplot2")
economics
```


## Line plot with `geom_line()`

```{r geom-line}
ggplot(economics, aes(date, unemploy)) + 
  geom_line()
```


## `geom_path()`

```{r geom-path}
ggplot(economics, aes(unemploy/pop, psavert)) +
  geom_path()
```


## `geom_step()`

```{r geom-step}
ggplot(economics, aes(unemploy/pop, psavert)) +
  geom_step()
```


## Drawing maps

* Drawing maps require the map data 

```{r}
world <- map_data("world")
world
```


## Drawing world map with `geom_polygon()`

```{r geom-polygon}
world <- map_data("world")
ggplot(world, aes(long, lat, group = group)) +
  geom_polygon()
```


## Other layers 

* There are more layers in `ggplot`!
* We won't cover the layers provided via functions starting with `stat_` but there are plenty of examples in the wild.
* There are also many extension packages to draw more plots: https://exts.ggplot2.tidyverse.org/gallery/ 
* It is normal to google and copy-and-paste code that you find on the web. 
* `ggplot2` is also [continually updated](https://github.com/tidyverse/ggplot2/blob/main/NEWS.md) so some new features may be added in the future. 






