---
subtitle: "Data Visualisation with R"
pagetitle: "Effective data visualistion"
author: "Emi Tanaka"
date: "28th November 2022"
host: "Australasian Applied Statistics Conference 2022"
institute: "Monash University"
strip-comments: true
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: true
    theme: assets/monash.scss
    show-slide-number: all
    controls: true
    width: 1280
    height: 720
    incremental: false
    auto-stretch: false
    css: [assets/custom.css, assets/pacman.css, assets/tachyons-addon.css]
    include-after: |
      <link rel="stylesheet" href="assets/syntax-highlight.css">
      <link rel="stylesheet" href="assets/fontawesome-free-6.1.1-web/css/all.min.css">
      <script src="assets/fontawesome-free-6.1.1-web/js/all.min.js"></script>
callout-icon: false
---




```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
library(colorspace)

df <- read_csv(here::here("data/census-birthplace.csv"))
df2021 <- df %>% 
  filter(census == 2021) %>% 
  filter(!birth %in% c("Total")) %>% 
  select(-census)
df2 <- df %>% 
  filter(census %in% c(2021, 2016)) %>% 
  filter(!birth %in% c("Total")) %>% 
  filter(!birth %in% c("Not Stated", "Other", "Australia")) %>% 
  group_by(census) %>% 
  mutate(rank = rank(-percentage)) %>% 
  filter(rank %in% 1:5) %>% 
  ungroup()

dfsex <- read_csv(here::here("data/census-birthplace-by-sex.csv")) %>% 
  filter(!birth %in% c("Total")) %>% 
  filter(!birth %in% c("Not Stated", "Other", "Australia")) %>% 
  mutate(birth = fct_reorder(birth, count, mean)) %>% 
  group_by(sex) %>% 
  mutate(rank = rank(-percentage)) %>% 
  filter(rank %in% 1:5) %>% 
  ungroup()


source("setup.R")
```


# Data story <i class="fas fa-book-open"></i> birth place, another look {background-color="#006DAE"}


## Another look {.scrollable}

::: callout-note

## Data Story 

India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found.

[-- The Guardian](https://www.theguardian.com/australia-news/2022/jun/28/india-now-third-most-common-place-of-birth-of-australian-residents-census-results-show)

:::

```{r}
df2 %>% 
  select(-rank) %>% 
  mutate(count = scales::comma(count, 1),
         percentage = scales::comma(percentage, 0.1)) %>% 
  knitr::kable(col.names = c("Birth place", "Count", "%", "Census Year"),
               align = "lrrr") %>% 
  kableExtra::kable_classic(full_width = FALSE)
```

## India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found

```{r census2-part1, fig.width = 8, fig.height = 3}
df2  %>% 
  mutate(birth = fct_reorder(birth, count, sum)) %>% 
  ggplot(aes(count, birth)) +
  geom_col() + 
  geom_col(data = ~filter(.x, birth=="India"),
           fill = "#006DAE") + 
  geom_text(aes(label = scales::percent(percentage/100, 0.1)),
            nudge_x = -70000, color = "white") +
  scale_x_continuous(labels = scales::comma) +
  theme(text = element_text(size = 18),
        axis.text.y = element_text(size = 8),
        plot.title.position = "plot") +
  facet_wrap(~census) + 
  scale_x_continuous(labels = scales::comma) +
  labs(y = "Birth place", x = "Number of Australian residents\n", caption = "Data source: Australian Census 2016 and 2021",
       title = "Top 5 countries of birth outside Australia")
```


. . . 

Does this show that India overtook China and New Zealand?




## India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found

```{r census2-part2, fig.width = 7, fig.height = 4}
df2 %>% 
  ggplot(aes(factor(census), count, color = birth)) +
  geom_line(aes(group = birth))  +
  theme(text = element_text(size = 18),
        axis.text.y = element_text(size = 8),
        plot.title.position = "plot") + 
  geom_point() + 
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Number of \nAustralian residents\n", caption = "Data source: Australian Census 2016 and 2021", x = "Census Year", color = "Birth place",
       title = "Top 5 countries of birth outside Australia")
```

. . . 

Should we show percentage instead of counts?


## India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found

```{r census2-part3, fig.width = 7, fig.height = 4}
df2 %>% 
  ggplot(aes(factor(census), percentage/100, color = birth)) +
  geom_line(aes(group = birth)) + 
  geom_point() + 
  theme(text = element_text(size = 18),
        axis.text.y = element_text(size = 8),
        plot.title.position = "plot") + 
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of\nAustralian residents\n", caption = "Data source: Australian Census 2016 and 2021", x = "Census Year", color = "Birth place",
       title = "Top 5 countries of birth outside Australia")
```


. . . 

The legend and the line order is different...


## India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found

```{r census2-part4, fig.width = 6.5, fig.height = 4}
df2 %>% 
  arrange(census, birth) %>% 
  mutate(birth = fct_reorder2(birth, census, percentage, last2)) %>% 
  ggplot(aes(factor(census), percentage/100, color = birth)) +
  geom_line(aes(group = birth)) + 
  geom_point() + 
  theme(text = element_text(size = 18),
        axis.text.y = element_text(size = 8),
        plot.title.position = "plot") + 
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of\nAustralian residents\n", caption = "Data source: Australian Census 2016 and 2021", x = "Census Year", color = "Birth place",
       title = "Top 5 countries of birth outside Australia")
```


. . . 

Maybe we can put the labels directly in the plot?

## India has overtaken China and New Zealand to become the third largest country of birth for Australian residents, 2021 census data has found {background-color="#D93F00"}

```{r census2-part5, dev.args=list(bg = "transparent")}
#| fig-width: 6.5
df2 %>% 
  arrange(census, birth) %>% 
  mutate(highlight = case_when(birth %in% c("India", "China", "New Zealand") ~ 1,
                               TRUE ~ 0.5)) %>% 
  ggplot(aes(factor(census), percentage/100, color = birth, alpha = factor(highlight))) +
  geom_line(aes(group = birth)) + 
  geom_point() + 
  scale_alpha_manual(values = c(0.5, 1)) +
  theme(text = element_text(size = 18, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        panel.grid.major = element_line(linetype = "dashed"),
        panel.grid.minor = element_line(linetype = "twodash"),
        axis.ticks.length = unit(0, "mm"),
        plot.title.position = "plot",
        legend.position = "none",
        plot.margin = margin(20, 20, 20, 20),
        plot.background = element_rect(fill = "#e3e3e3",
                                       color = "black",
                                       linewidth = 1),
        panel.background = element_blank()) + 
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(label = birth), 
            data = ~{ .x %>% 
                filter(census==2021) %>% 
                mutate(percentage = case_when(birth=="China" ~ percentage + 0.05,
                                              birth=="New Zealand" ~ percentage - 0.05,
                                              TRUE ~ percentage))}, 
            hjust = 0, nudge_x = 0.05) +
  labs(y = "Percentage of\nAustralian residents\n", caption = "Data source: Australian Census 2016 and 2021", x = "Census Year", color = "Birth place",
       title = "Top 5 countries of birth outside Australia")
```

```{css, echo = FALSE}
.reveal h2 {
  font-size: 34px!important;
}
```
