---
subtitle: "Data Visualisation with R"
pagetitle: "Color choices"
author: "Emi Tanaka"
date: "28th November 2022"
host: "Australasian Applied Statistics Conference 2022"
institute: "Monash University"
strip-comments: true
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: true
    theme: assets/monash.scss
    show-slide-number: all
    controls: true
    width: 1280
    height: 720
    incremental: true
    auto-stretch: false
    css: [assets/custom.css, assets/pacman.css, assets/tachyons-addon.css]
    include-after: |
      <link rel="stylesheet" href="assets/syntax-highlight.css">
      <link rel="stylesheet" href="assets/fontawesome-free-6.1.1-web/css/all.min.css">
      <script src="assets/fontawesome-free-6.1.1-web/js/all.min.js"></script>
callout-icon: false
execute:
  echo: true
---


## {#title-slide background-image="images/bg11.jpeg"}


```{r}
#| label: setup
#| include: false
library(tidyverse)
source("setup.R")
```

```{r}
#| child: title-slide.qmd
```


# Color space

::: {.f3 .monash-gray80}

Zeileis, Fisher, Hornik, Ihaka, McWhite,
Murrell, Stauffer, Wilke (2019). colorspace: A
Toolbox for Manipulating and Assessing Colors and
Palettes. *arXiv 1903.06490*

Zeileis, Hornik, Murrell (2009). Escaping RGBland:
Selecting Colors for Statistical Graphics.
_Computational Statistics & Data Analysis_ 53(9)
3259-3270

:::

## Qualitative palettes

* Designed for categorical variable with no particular ordering 


```{r, fig.height = 4, fig.width = 8, echo = TRUE}
#| code-line-numbers: false
colorspace::hcl_palettes("Qualitative", plot = TRUE, n = 7)
```

## Sequential palettes

* Designed for ordered categorical variable or number going from low to high (or vice-versa) 


```{r, fig.height = 4, fig.width = 10, echo = TRUE}
#| code-line-numbers: false
colorspace::hcl_palettes("Sequential", plot = TRUE, n = 7)
```


## Diverging palettes

* Designed for ordered categorical variable or number going from low to high (or vice-versa) with a neutral value in between


```{r, fig.height = 6, fig.width = 8, echo = TRUE}
#| code-line-numbers: false
colorspace::hcl_palettes("Diverging", plot = TRUE, n = 7)
```

## Colorblindness 

Colorblindness affect roughly 1 in 8 men.

```{r, fig.width = 12}
colorblindr::cvd_grid(gprox)
```

Check your color choices using the [`colorblindr`](https://github.com/clauswilke/colorblindr) package or otherwise. 

