---
subtitle: "Data Visualisation with R"
pagetitle: "Effective data visualistion"
author: "Emi Tanaka"
date: "28th November 2022"
host: "Australasian Applied Statistics Conference 2022"
institute: "Monash University"
strip-comments: true
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: true
    theme: assets/monash.scss
    show-slide-number: all
    controls: true
    width: 1280
    height: 720
    incremental: false
    auto-stretch: false
    css: [assets/custom.css, assets/pacman.css, assets/tachyons-addon.css]
    include-after: |
      <link rel="stylesheet" href="assets/syntax-highlight.css">
      <link rel="stylesheet" href="assets/fontawesome-free-6.1.1-web/css/all.min.css">
      <script src="assets/fontawesome-free-6.1.1-web/js/all.min.js"></script>
callout-icon: false
---




```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
library(colorspace)

df <- read_csv(here::here("data/census-birthplace.csv"))
df2021 <- df %>% 
  filter(census == 2021) %>% 
  filter(!birth %in% c("Total")) %>% 
  select(-census)
df2 <- df %>% 
  filter(census %in% c(2021, 2016)) %>% 
  filter(!birth %in% c("Total")) %>% 
  filter(!birth %in% c("Not Stated", "Other", "Australia")) %>% 
  group_by(census) %>% 
  mutate(rank = rank(-percentage)) %>% 
  filter(rank %in% 1:5) %>% 
  ungroup()

dfsex <- read_csv(here::here("data/census-birthplace-by-sex.csv")) %>% 
  filter(!birth %in% c("Total")) %>% 
  filter(!birth %in% c("Not Stated", "Other", "Australia")) %>% 
  mutate(birth = fct_reorder(birth, count, mean)) %>% 
  group_by(sex) %>% 
  mutate(rank = rank(-percentage)) %>% 
  filter(rank %in% 1:5) %>% 
  ungroup()


source("setup.R")
```


# Data story <i class="fas fa-book-open"></i> birth place {background-color="#006DAE"}



## Birth place from the 2021 Australian Census {.scrollable }

```{r}
df2021 %>% 
  mutate(count = scales::comma(count, 1),
         percentage = scales::comma(percentage, 0.1)) %>% 
  knitr::kable(col.names = c("Birth place", "Count", "%"),
               align = "lrr") %>% 
  kableExtra::kable_classic(full_width = FALSE)
```

## Birth place from the 2021 Australian Census 

```{r census2021-part1, fig.width = 12, fig.height = 4}
df2021 %>% 
  ggplot(aes(birth, percentage)) +
  geom_col() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 45, size = 10, hjust = 1)) + 
  labs(x = "", y = "Percentage", caption = "Data source: Australian Census 2021")
```

. . . 

Which birth place is the third largest among people in Australia?


## Birth place from the 2021 Australian Census 

```{r census2021-part2, fig.width = 12, fig.height = 4}
df2021 %>% 
  mutate(birth = fct_reorder(birth, -percentage)) %>% 
  ggplot(aes(birth, percentage)) +
  geom_col() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 45, size = 10, hjust = 1)) + 
  labs(x = "", y = "Percentage", caption = "Data source: Australian Census 2021")
```

. . . 

<i class="fa-solid fa-face-meh" style=" transform: rotate(-45deg)"></i> Can you read the labels without tilting your head?

## Birth place from the 2021 Australian Census { .scrollable}

```{r census2021-part4, fig.height = 12, fig.width = 4}
df2021 %>% 
  mutate(birth = fct_reorder(birth, percentage)) %>% 
  ggplot(aes(percentage, birth)) +
  geom_col() +
  theme(text = element_text(size = 18),
        axis.text.y = element_text(size = 8)) + 
  labs(y = "", x = "Percentage", caption = "Data source: Australian Census 2021")
```

. . . 

What's the data story? 

## India now third most common place of birth of Australian residents, census results show {background-color="#D93F00"}

::: flex

::: w-60

```{r census2021-part5, dev.args=list(bg = "transparent")}
total2021 <- df %>% 
  filter(birth=="Total" & census==2021) %>% 
  pull(count)
auperc2021 <- df %>% 
  filter(birth=="Australia" & census==2021) %>% 
  pull(percentage)
nsperc2021 <- df %>% 
  filter(birth=="Not Stated" & census==2021) %>% 
  pull(percentage)
df2021 %>% 
  arrange(desc(percentage)) %>% 
  filter(!birth %in% c("Australia", "Other", "Not Stated")) %>% 
  slice(1:5) %>% 
  mutate(birth = fct_reorder(birth, count)) %>% 
  ggplot(aes(count, birth)) +
  geom_col() +
  geom_col(data = ~filter(.x, birth=="India"),
           fill = "#006DAE") + 
  geom_text(aes(label = scales::percent(percentage/100, 0.1)),
            nudge_x = -50000, color = "white") +
  scale_x_continuous(labels = scales::comma, expand  = c(0, 0)) +
  theme(text = element_text(size = 18, color = "white"),
        axis.text = element_text(size = 8, color = "white"),
        axis.ticks.length = unit(0, "mm"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(linetype = "dashed"),
        panel.grid.minor.x = element_blank(),
        plot.title.position = "plot",
        plot.background = element_rect(fill = "transparent",
                                       color = "transparent"),
        panel.background = element_blank()) + 
  labs(y = "Birth place", x = "Number of Australian residents\n", caption = "Data source: Australian Census 2021",
       title = "Top 5 countries of birth outside Australia")
```


:::

::: {.w-40 .f3}

<br><br><br>

* The text on the bar shows the percentage out of `r scales::comma(total2021)` Australian residents born in that place.  

* There were `r scales::percent(nsperc2021/100, 0.1)` of Australian residents who did not state their birth place.

* The top country of birth place is Australia with  `r scales::percent(auperc2021/100, 0.1)` of Australian residents born in Australia.


<i class="fa-solid fa-newspaper"></i> Story from [The Guardian](https://www.theguardian.com/australia-news/2022/jun/28/india-now-third-most-common-place-of-birth-of-australian-residents-census-results-show).

:::


:::



```{css, echo = FALSE}
.reveal h2 {
  font-size: 34px!important;
}
```
