---
title: "Exercise 1: R Graphics"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Check out your knowledge in R Graphics.
---

```{r setup, child="setup.Rmd", eval = file.exists("setup.Rmd")}
```

## Welcome

::: aim

The aim of this exercise is to:

- have a check of your knowledge of R graphics system,
- see if this exercise is working for you, and 
- briefly review base plots/graphics.

:::

## Check your knowledge


```{r check1, echo=FALSE}
question("Which of the following is a contributed R package?",
  answer("ggplot2", correct = TRUE),
  answer("grid"),
  answer("graphics"),
  answer("grDevices")
)
```

```{r check2, echo=FALSE}
question("Which R package actually renders the graphics in R?",
  answer("ggplot2"),
  answer("grid"),
  answer("graphics"),
  answer("grDevices", correct = TRUE)
)
```

```{r check3, echo=FALSE}
question("There are two primary graphic models in R: the base and grid graphics. Which one is ggplot2 using?",
  answer("ggplot2"),
  answer("grid", correct = TRUE),
  answer("graphics"),
  answer("grDevices")
)
```

## Base plots

When we say "base plots" or "base graphics", these refer to plots created using the `graphics` package. The `graphics` package is a core package that is already loaded when you launch a standard installation of R.

Let's have a look at the `cars` dataset that contains two variables `speed` and `dist`. Below shows the structure of the `cars` dataset.

```{r}
str(cars)
```

Base plots are not the focus of this workshop, but below are some examples. 

### Scatter plot


Try running the code to see if the system is working for you. You should see a scatter plot. Try also uncommenting out the second line to see the line of best fit using least squares approach.

```{r base1, exercise = TRUE}
plot(dist ~ speed, data = cars)
# abline(coef(lm(dist ~ speed, data = cars)))
```


### Histogram

Next we draw the histogram:

```{r base2, exercise = TRUE}
hist(cars$dist)
```



### Boxplot

Can you how to draw a boxplot for the variable `dist` in the `cars` dataset? Try checking the solution if you are not sure.

```{r base3, exercise = TRUE}
___(cars$dist)
```

```{r base3-solution}
boxplot(cars$dist)
```

