---
title: "Exercise 2: Getting started with ggplot2"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, child="setup.Rmd", eval = file.exists("setup.Rmd")}
```

## Objective  {data-progressive=FALSE}

::: aim

You are going to examine two datasets, `diamonds` from `ggplot2` package and `crampton.pig` from `agridat` package, by making targetted plots. You can find the structure of these datasets below. Enter `?agridat::crampton.pig` or `?ggplot2::diamonds` in R for the documentations of the dataset. 

:::


```{r pkgs, message = FALSE, warning = FALSE}
library(tidyverse) # contains ggplot2
library(agridat) # for `crampton.pig` data
```

### Data: `crampton.pig`

```{r pig-data}
str(crampton.pig)
```

### Data: `diamonds` 

```{r diamonds-data}
str(diamonds)
```


## Target plots for `crampton.pig`

### Exercise 1

```{r ex01s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(feed)) + 
  geom_density() 
```

```{r ex01, exercise = TRUE}
ggplot(crampton.pig, aes(___)) + 
  geom_density() 
```

```{r ex01-solution, ref.label="ex01s"}
```

### Exercise 2

Now try changing the y-axis so it shows `count` instead of `density`.

```{r ex02s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(feed)) + 
  geom_density(aes(y = after_stat(count))) 
```

```{r ex02, exercise = TRUE}
ggplot(crampton.pig, aes(___)) + 
  geom____(aes(y = ___)) 
```

```{r ex02-solution, ref.label="ex02s"}
```

### Exercise 3

```{r ex03s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(feed)) + 
  geom_boxplot() 
```

```{r ex03, exercise = TRUE}
ggplot(crampton.pig, aes(___)) + 
  geom____() 
```

```{r ex03-solution, ref.label="ex03s"}
```


### Exercise 4

```{r ex04s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(weight1, weight2, color = treatment, size = feed)) + 
  geom_point()
```

```{r ex04, exercise = TRUE}
ggplot(crampton.pig, aes(___, ___, color = ___, size = ___)) + 
  geom____()
```

```{r ex04-solution, ref.label="ex04s"}
```

### Exercise 5

```{r ex05s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(rep, treatment, fill = weight2 - weight1)) + 
  geom_tile(color = "black", linewidth = 2)
```

```{r ex05, exercise = TRUE}
ggplot(crampton.pig, aes(___, ___, fill = ___)) + 
  geom_tile(color = "black", linewidth = 2)
```

```{r ex05-solution, ref.label="ex05s"}
```

### Exercise 6

```{r ex06s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(weight2 - weight1, fill = treatment)) + 
  geom_density(alpha = 0.7) 
```

```{r ex06, exercise = TRUE}
ggplot(crampton.pig, aes(___, fill = ___)) + 
  geom_density(alpha = 0.7) 
```

```{r ex06-solution, ref.label="ex06s"}
```


## Target plots for `diamonds`


### Exercise 7

```{r ex07s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(price)) + 
  geom_histogram() 
```

```{r ex07, exercise = TRUE}
ggplot(diamonds, aes(___)) + 
  geom____() 
```

```{r ex07-solution, ref.label="ex07s"}
```

### Exercise 8

Try changing the bin width to see how the histogram changes. 

```{r ex08s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(price)) + 
  geom_histogram(binwidth = 1) 
```

```{r ex08, exercise = TRUE}
ggplot(diamonds, aes(___)) + 
  geom____(___ = 1) 
```

```{r ex08-solution, ref.label="ex08s"}
```

### Exercise 9

```{r ex09s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(price)) + 
  stat_bin(geom = "line")
```

```{r ex09, exercise = TRUE}
ggplot(diamonds, aes(___)) + 
  stat____(___ = "line") 
```

```{r ex09-solution, ref.label="ex09s"}
```



### Exercise 10

```{r ex10s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(clarity, price, fill = cut)) + 
  geom_boxplot()
```

```{r ex10, exercise = TRUE}
ggplot(diamonds, aes(___, ___, fill = ___)) + 
  geom____()
```

```{r ex10-solution, ref.label="ex10s"}
```


### Exercise 11

```{r ex11s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(carat, price)) + 
  geom_density_2d_filled()
```

```{r ex11, exercise = TRUE}
ggplot(heart2, aes(___, ___)) + 
  geom_density_2d_filled()
```

```{r ex11-solution, ref.label="ex11s"}
```


### Exercise 12

```{r ex12s, echo = FALSE, class = "target", message = FALSE}
ggplot(diamonds, aes(carat, price)) + 
  geom_hex()
```

```{r ex12, exercise = TRUE}
ggplot(diamonds, aes(___, ___)) + 
  geom____()
```

```{r ex12-solution, ref.label="ex12s"}
```



This is the last question! Well done getting to here!
