---
title: "Exercise 4: Barplots and position adjustments in ggplot2"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, child="setup.Rmd", eval = file.exists("setup.Rmd")}
```

## Objective  {data-progressive=FALSE}

::: aim

You are going to use two datasets, `crampton.pig` and `reid.grasses` data from the `agridat` package in order to study how the position adjustment and coordinate systems work. The structure of these datasets are shown below.

:::


```{r pkgs, message = FALSE, warning = FALSE}
library(tidyverse) # contains ggplot2
library(agridat)
```

### Data: `crampton.pig`

```{r pig-data}
str(crampton.pig)
```

### Data: `reid.grasses`

```{r reid.grasses}
str(reid.grasses)
```

## Exercise 1

```{r ex01s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(treatment, feed, fill = rep)) + 
  geom_col(position = "dodge", color = "black")
```

```{r ex01, exercise = TRUE}
ggplot(crampton.pig, aes(___, ___, fill = ___)) + 
  geom_col(position = "dodge", color = "black")
```

```{r ex01-solution, ref.label="ex01s"}
```

## Exercise 2

The plot below is called a "rose plot". 

```{r ex02s, echo = FALSE, class = "target"}
ggplot(crampton.pig, aes(treatment, feed, fill = rep)) + 
  geom_col(position = "dodge") + 
  coord_polar("x")
```

Try making this rose plot now.

```{r ex02, exercise = TRUE}
ggplot(crampton.pig, aes(___, ___, fill = ___)) + 
  geom____(___) + 
  coord____(___)
```

```{r ex02-solution, ref.label="ex02s"}
```

## Exercise 3

Position adjustments are also applicable beyond barplots like below.

```{r ex03s, echo = FALSE, class = "target", message = FALSE, fig.height = 6}
ggplot(reid.grasses, aes(gen, drymatter, fill = year)) + 
  geom_dotplot(position = "dodge",
               binaxis = "y",
               stackdir = "center") + 
  coord_flip()
```

```{r ex03, exercise = TRUE}
ggplot(reid.grasses, aes(___, ___, fill = ___)) + 
  geom_dotplot(position = ___,
               binaxis = "y",
               stackdir = "center") + 
  coord____()
```

```{r ex03-solution, ref.label="ex03s", fig.height = 6}
```
