---
title: "Exercise 3: Multiple layers"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, child="setup.Rmd", eval = file.exists("setup.Rmd")}
```

## Objective  {data-progressive=FALSE}

::: aim

* You are going to examine the datasets, `nass.wheat`, `nass.corn`, and `nass.soybean` data from the `agridat` package. The structure of these datasets are shown below.

* In this exercise, you will need to wrangle the data in the right format (which may be achieved by multiple approaches) before plotting. 

:::

```{r, include = FALSE}
library(maps)
```


```{r pkgs, message = FALSE, warning = FALSE}
library(tidyverse) # contains ggplot2
library(agridat) # contains the datasets
```

### Data: `nass.wheat`

```{r}
str(nass.wheat)
```

### Data: `nass.corn`

```{r}
str(nass.corn)
```

### Data: `nass.soybean`

```{r}
str(nass.soybean)
```

## Exercise 1

```{r ex01s, echo = FALSE, class = "target", message = FALSE, warning = FALSE}
wheat_corn_data <- left_join(nass.wheat, 
                             nass.corn, 
                             suffix = c("_wheat", "_corn"),
                             by = c("year", "state"))

ggplot(wheat_corn_data, aes(yield_corn, yield_wheat)) + 
  geom_point(aes(color = year)) +
  geom_density_2d(color = "black") 
```

```{r ex01, exercise = TRUE}
wheat_corn_data <- ___

ggplot(wheat_corn_data, aes(___, ___)) + 
  geom____(aes(color = ___)) +
  geom_density_2d(color = "black") 
```

```{r ex01-hint}
wheat_corn_data <- left_join(nass.wheat, 
                             nass.corn, 
                             suffix = c("_wheat", "_corn"),
                             by = c("year", "state"))
```

```{r ex01-solution, ref.label="ex01s"}
```

## Exercise 2

```{r ex02s, echo = FALSE, class = "target"}
crops_data <- bind_rows(mutate(nass.wheat, crop = "Wheat"),
                        mutate(nass.corn, crop = "Corn"),
                        mutate(nass.soybean, crop = "Soy")) 

ggplot(crops_data, aes(crop, yield)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1)
```

```{r ex02, exercise = TRUE}
crops_data <- ___

ggplot(crops_data, aes(___, ___)) + 
  geom____() + 
  geom____(width = 0.1)
```

```{r ex02-hint}
crops_data <- bind_rows(mutate(nass.wheat, crop = "Wheat"),
                        mutate(nass.corn, crop = "Corn"),
                        mutate(nass.soybean, crop = "Soy")) 
```

```{r ex02-solution, ref.label="ex02s"}
```


## Exercise 3

```{r ex03s, echo = FALSE, class = "target"}
wheat2011_data <- nass.wheat %>% 
  filter(year == 2011) 

highlight <- function(.data) subset(.data, acres > 4000000 | yield > 80)

ggplot(wheat2011_data, aes(acres, yield)) + 
  geom_point() + 
  geom_point(data = highlight, color = "red") + 
  geom_text(data = highlight, aes(label = state), nudge_y = 5) 
```

```{r ex03, exercise = TRUE}
wheat2011_data <- nass.wheat %>% 
  ___(year == 2011) 

highlight <- function(.data) subset(.data, acres > 4000000 | yield > 80)

ggplot(wheat2011_data, aes(___, yield)) + 
  ___() + 
  ___(data = ___, color = "red") + 
  geom_text(data = ___, aes(label = ___), nudge_y = 5) 
```

```{r ex03-solution, ref.label="ex03s"}
```

## Exercise 4

Let's draw a choropleth US map of the wheat yield in 2011. 

```{r ex04s, echo = FALSE, class = "target"}
wheat2011_data_with_map <- nass.wheat %>% 
  filter(year == 2011) %>% 
  mutate(region = tolower(state)) %>% 
  right_join(map_data("state"), by = "region")

ggplot(wheat2011_data_with_map, aes(long, lat)) + 
  geom_polygon(aes(group = group, fill = yield), color = "black") +
  annotate("label", x = -120, y = 27, label = "2011", size = 8)
```

```{r ex04-hint}
wheat2011_data_with_map <- nass.wheat %>% 
  filter(year == 2011) %>% 
  mutate(region = tolower(state)) %>% 
  right_join(map_data("state"), by = "region")
```


```{r ex04, exercise = TRUE}
wheat2011_data_with_map <- nass.wheat %>% 
  ___

ggplot(wheat2011_data_with_map, aes(___, ___)) + 
  geom_polygon(aes(___), color = "black") +
  ___(___, x = -120, y = 27, label = "2011", size = 8)
```

```{r ex04-solution, ref.label="ex04s"}
```


## Exercise 5

Consider fitting a model to the log of the yield of wheat (with base 10) with year as a covariate as shown below in the first line of code. Try constructing a residual plot based on this fit. What do you think of the fit of the model?

```{r ex05s, echo = FALSE, class = "target"}
fit <- lm(log10(yield) ~ year, data = nass.wheat)

ggplot(nass.wheat, aes(year, residuals(fit))) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey")
```

```{r ex05, exercise = TRUE}
fit <- lm(log10(yield) ~ year, data = nass.wheat)

ggplot(nass.wheat, aes(year, residuals(fit))) + 
  ___ + 
  geom____(yintercept = 0, linetype = "dashed", color = "grey")
```

```{r ex05-solution, ref.label="ex05s"}
```
